CYO Project Report - Delivery Prediction System

### 2021-07-20 finalReportPrep.Rmd
fine tuned files

== TODO additive smoothing/Bayes estimator with target encoding

@@

However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied **target encoding** on the product categories attribute.

But to do that, first, we checked number of observations per product category and noticed that the distribution could require **regularization** to prevent overfitting. Secondly, target encoding could be erroneous if done against the random dataset.

To solve these, we identified the regularization weight to use by carrying out target encoding for different datasets set of weights, selecting single weight per dataset based on performance metric i.e. RMSE and then we took a mean of weights identified across the datasets.

Once weight was identified, to avoid any randomness due to choice of the dataset, we computed target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, seller ids similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, which are a lot. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for seller ids we applied target encoding and followed strategy used for product categories.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer states similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_state) %>% nrow()`** customer states. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied target encoding and followed strategy used for product categories.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer locations similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_loc) %>% nrow()`** customer locations. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied target encoding and followed strategy used for product categories.

@@

In addition, as earlier shared, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting. Also, any undesired features were removed, reducing train_subset and test_subset to only the desired features.

@@

In addition, categorical features were converted to numeric form with the help of target encoding with a weight term computed earlier for purpose of regularization and to avoid overfitting. Also, any undesired features were removed, reducing training_dataset and validation_dataset to only the desired features.

@@

During the course of project, we carried out data cleaning, identified features and derived features from existing information available. In addition, converted categorical features to numeric form and applied various techniques like target encoding, scaling and normalization as required and regularized the required features.

>>>>

@@

However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied additive smoothing based **target encoding** on the product categories attribute.

Additive smoothing based target encoding for a categorical variable = $(global\_mean * weight + N * u_t,c)/(weight+N)$

where $u_{t,c}$ is mean of target/outcome for a given category and weight is the smoothing factor or amount of regularization

But to do that, first we checked number of observations per product category, to understand the range of weights to consider. The target encoding could be erroneous if done with larger weight or against the random dataset.

To solve these, we identified the smoothing factor(or weight) to use by carrying out target encoding for different datasets for a set of weights, selecting single weight per dataset based on performance metric i.e. RMSE and then we took a mean of weights identified across the datasets.

Once weight was identified, to avoid any randomness due to choice of the dataset, we computed additive smoothing based target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, seller ids similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, which are a lot. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for seller ids we applied additive smoothing based target encoding and followed strategy used for product categories.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer states similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_state) %>% nrow()`** customer states. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied additive smoothing based target encoding and followed strategy used for product categories.

@@

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer locations similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_loc) %>% nrow()`** customer locations. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied additive smoothing based target encoding and followed strategy used for product categories.

@@

In addition, as earlier shared, categorical features were converted to numeric form with the help of additive smoothing based target encoding with a smoothing factor/weight term computed earlier to avoid overfitting. Also, any undesired features were removed, reducing train_subset and test_subset to only the desired features.

@@

In addition, categorical features were converted to numeric form with the help of additive smoothing based target encoding with a weight term computed earlier for purpose of smoothing and to avoid overfitting. Also, any undesired features were removed, reducing training_dataset and validation_dataset to only the desired features.

@@

During the course of project, we carried out data cleaning, identified features and derived features from existing information available. In addition, converted categorical features to numeric form and applied various techniques like additive smoothing based target encoding, scaling and normalization as required.
